### 2019-02-22 report/final-report.tex
added some experimental results

Reffing a \Cref{fig:eg}.

\begin{figure}
  a picture is here
  \caption{An example figure}
  \label{fig:eg}
\end{figure}

>>>>

% From bijective lens paper

% \newcommand{\SOptician}{Optician\textsubscript{S}}
% \newcommand{\SynthSymLens}{\PCF{SynthSymLens}\xspace}
% \newcommand{\RXSearchState}{\ensuremath{\mathit{pq}}}
% \newcommand{\ToStochastic}{\PCF{ToStochastic}\xspace}
% \newcommand{\LC}{\ensuremath{\mathit{lc}}}
% \newcommand{\Best}{\ensuremath{\mathit{best}}}
% \newcommand{\RXSearch}{\PCF{RXSearch}\xspace}
% \newcommand{\Continue}{\PCF{Continue}\xspace}
% \newcommand{\PQ}{\PCF{PQ}\xspace}
% \newcommand{\SynthDNFLens}{\PCF{SynthDNFLens}}
% \newcommand{\ToLens}{\ensuremath{\Uparrow}}
% \newcommand{\ToLensOf}[1]{\ensuremath{\ToLens{}\mkern-4mu #1}}
% \newcommand{\ToDNFRegexText}{\PCF{ToDNFRegex}}
% \newcommand{\Beautify}{\PCF{Beautify}}
% \newcommand{\RigidSynth}{\PCF{RigidSynth}}
% \newcommand{\GreedySynth}{\PCF{GreedySynth}\xspace}
% \newcommand{\RigidSynthInternal}{\PCF{RigidSynthInternal}}
% \newcommand{\RigidSynthSequence}{\PCF{RigidSynthSeq}}
% \newcommand{\RigidSynthAtom}{\PCF{RigidSynthAtom}}
% \newcommand{\GetDNFNormalizer}{\PCF{GetDNFNormalizer}}
% \newcommand{\CreatePQueue}{\PCF{CreatePQueue}}
% \newcommand{\GetTransitiveSet}{\PCF{GetTransitiveSet}}
% \newcommand{\GetCurrentSet}{\PCF{GetCurrentSet}}
% \newcommand{\Pop}{\PCF{Pop}}
% \newcommand{\ExpandOnce}{\PCF{ExpandOnce}}
% \newcommand{\ExpandRequired}{\PCF{ExpandRequired}}
% \newcommand{\FixProblemElts}{\PCF{FixProblemElts}}
% \newcommand{\Expand}{\PCF{Expand}\xspace}
% \newcommand{\ForceExpand}{\PCF{ForceExpand}}
% \newcommand{\Reveal}{\PCF{Reveal}}
% \newcommand{\Map}{\PCF{Map}}
% \newcommand{\EnqueueMany}{\PCF{EnqueueMany}}
% \newcommand{\ReturnVal}[1]{\ensuremath{\Return\,#1}}
% \newcommand{\CurrentSet}{\ensuremath{\mathit{CS}}}
% \newcommand{\TransitiveSet}{\ensuremath{\mathit{TS}}}

% \newcommand{\SSOpt}{\ensuremath{\mathbf{SS}}}
% \newcommand{\SSNCOpt}{\ensuremath{\mathbf{SSNC}}}
% \newcommand{\BSOpt}{\ensuremath{\mathbf{BS}}}
% \newcommand{\BSNCOpt}{\ensuremath{\mathbf{BSNC}}}
% \newcommand{\AnyOpt}{\ensuremath{\mathbf{Any}}}
% \newcommand{\FLOpt}{\ensuremath{\mathbf{FL}}}
% \newcommand{\CCOpt}{\ensuremath{\mathbf{DC}}}
% \newcommand{\NSOpt}{\textbf{NS}}
% \newcommand{\NROpt}{\textbf{NR}}

We have implemented our synthesis algorithms in
of OCaml and integrated them into Boomerang system,
where they are available as open source code~\cite{GitHub}.
Users of Boomerang can write lenses by hand, specify them as
synthesis tasks, or do a combination of both.

We have analyzed our algorithms theoretically, proving various
soundness and completeness results.  We have also analyzed our
algorithms empirically, studying their impact on a range of real and
synthetic benchmarks.  Our full results have been published in a
series of conference papers and technical
reports~\cite{bijective,quotient,symmetric}.  In the rest of this
section, we summarize some of the key results.

\subsection{Bijective Synthesis Algorithms}

We first study the effectiveness of the pure bijective synthesis
algorithm by evaluating its performance on a set of 39
benchmark programs.

All evaluations were performed on a 2.5 GHz Intel Core i7 processor with 16 GB
of 1600 MHz DDR3 running macOS Sierra.

\paragraph*{Benchmark Suite Construction}
We constructed our benchmarks primarily by adapting examples from
Augeas~\cite{augeas} and 
Flash Fill~\cite{gulwani-popl-2014}.

Augeas is a configuration editing system for Linux that uses lens
combinators similar to those in Boomerang. However, it transforms
strings on the left to structured trees on the right rather than
transforming strings to strings.
We adapted these Augeas lenses to our setting by converting the
right-hand sides to strings that correspond to serialized versions
of the tree formats.  
We derived 29 of the benchmark tests by
adapting the first 27 lenses in alphabetical order, as well as the lenses
\CF{aug/xml-firstlevel} and \CF{aug/xml} that were referenced
by the `A' lenses.
Furthermore, the 12 last synthesis problems derived
from Augeas were tested after \Optician{} was
finalized, demonstrating that the optimizations were not
overtuned to perform well on the testing data.

Flash Fill is a system that allows users to specify common string
transformations by example~\cite{gulwani-popl-2014}.  
We derived three benchmarks from the first few examples in the
paper and one from the running example on
extracting phone numbers.

Both Augeas and Flash Fill permit non-bijective transformations.
To test our system on these benchmarks, we had to convert them into
bijective transformations by hand.  More specifically, in many of the
benchmarks, we normalized
the whitespace in the documents so the amount of whitespace was equal
in the source and target.  In a few of the benchmarks, either the
source or the target was missing information present in the other
format.  We modified those benchmarks by adding the missing
information back into the file.

Finally, we added custom examples to highlight weaknesses of
our algorithm (\CF{cap-prob} and \CF{2-cap-prob}) 
and to test situations for which we thought the tool would be
particularly useful (\CF{workitem-probs}, \CF{date-probs}, \CF{bib-prob},
and \CF{addr-probs}).   These examples convert between work item formats, date
formats, bibliography formats, and address formats, respectively.

\begin{figure}
  \centering
  \begin{subfigure}[b]{.49\textwidth}
    \centering
    \includegraphics{figs/specsizes}
    \caption{}
    \label{subfig:lenssize}
  \end{subfigure}
  \begin{subfigure}[b]{.49\textwidth}
    \includegraphics{figs/examplesused}
    \caption{}
    \label{subfig:examplesused}
  \end{subfigure}
  \caption{Sizes of Specifications.
    In (a), we show how many benchmarks are defined in our suite using a
    given number of AST nodes or fewer.
    In (b), we show how many benchmarks are defined in our suite using a
    given number of examples or fewer.}
  \label{fig:definition-sizes}
\end{figure}

Figure~\ref{fig:definition-sizes} shows the complexity of our type-based
specifications as well as our example counts.
An average benchmark has a type-based specification that can be represented using
310 AST node, and requires 1.1 input/output examples.
Our benchmarks vary from simple problems, like changing the
representation of dates
(with a specification size of 85, and a generated lens size of 79), to 
complex tasks, like transforming configuration files for server monitoring
software into dictionary form (with a specification size of 670 and
a generated lens size of 651).  On average, the size of the generated lens is
89\% the size of its type specifications.

\paragraph*{Importance of Examples}

\begin{figure}
  \centering
  \includegraphics{figs/examples.eps}
  \caption{Average number of random examples required to synthesize benchmark
    programs.  {\bf Experimental Average} is the average number of randomly
    generated examples needed to correctly synthesize the lens.  {\bf
      Determinize Permutations} is the theoretical number of examples required
    to determinize the choice all the permutations in \RigidSynth{}.
    In practice, far fewer examples are
    needed to synthesize the correct lens than would be predicted by the number
    required to determinize permutations.}
  \label{fig:exs-reqd}
\end{figure}

To evaluate how many user-supplied examples the algorithm requires in
practice, we \textit{randomly} generated appropriate source/target
pairs, mimicking what a na\"{i}ve user might do.  We did not write the
examples by hand out of concern that our knowledge of the synthesis
algorithm might bias the selection. Figure~\ref{fig:exs-reqd} shows
the number of randomly generated examples it takes to synthesize the
correct lens averaged over ten runs.  The synthesis algorithm almost never needs
any examples: only 5 benchmarks need a nonzero number of examples to
synthesize the correct lens and only one, \CF{cust/workitem-probs} required over
10 randomly generated examples.
A clever user may be able to reduce the
number of examples further by selecting examples carefully; we
synthesized \CF{cust/workitem-probs} with only 8 examples.

These numbers are low because there are relatively few well-typed
bijective lenses between any two source and target regular expressions. 
As one would expect, the benchmarks where there are multiple ways to
map source data to the target (and vice versa) require the most examples.
For example, the benchmark \CF{cust/workitem-probs} requires a large number of
examples because it
must differentiate between data in different text fields in both the
source and target and map between them appropriately.  As these text fields are
heavily permuted
(the legacy format ordered fields by a numeric ID, where
the modern format ordered fields alphabetically) and fields can be
omitted, a number of examples are needed to correctly identify the mapping
between fields.

The average number of examples to
infer the correct lens does not tell the whole story.  The system will
stop as soon as it finds a well typed lens that satisfies the supplied examples.
This inferred lens may or may not 
correctly handle unseen examples that correspond to
unexercised portions of the source and target regular expressions.
Figure~\ref{fig:exs-reqd} lists
the number of examples that are required to determinize the generation of
permutations in \RigidSynth{}.
Intuitively, this number represents the maximum number of
examples that a user must supply to guide the synthesis engine if it
always guesses the wrong permutation when multiple permutations can be used to
satisfy the specification.

The average number of examples is so much lower than the maximum
number of required examples because of correspondences in how we wrote
the regular expressions for the source and target data formats. 
Specifically, when we had corresponding disjunctions in both the
source and the target, we ordered them the same way.  The algorithm
uses the supplied ordering to guide its search, and so the system
requires fewer examples.   We did not write the examples in this style
to facilitate synthesis, but rather because maintaining similar
subparts in similar orderings makes the types much easier to 
read. We expect that most users would do the same.

\paragraph*{Comparison Against Other Tools}
%
We are the first tool to synthesize bidirectional transformations between data
formats, so there is no tool to which we can make an apple-to-apples comparison.
Instead, we compare against tools for generating unidirectional
transformations instead. 
Figure~\ref{fig:synthesis-times} includes a comparison against two other
well-known tools that synthesize
text transformation and extraction functions from examples -- Flash Fill and FlashExtract.  For this
evaluation, we used the version of these tools distributed through the
PROSE project~\cite{prose}.

\begin{figure}
  \includegraphics{figs/times}
  \caption{
    Number of benchmarks that can be solved by a given algorithm in a given
    amount of time.
    \Optician{} is our bijective synthesis algorithm including a set
    of optimizations we have implemented.
    \FlashExtractMode{} is the existing FlashExtract system.  \FlashFillMode{} is
    the existing Flash Fill system.  \NaiveMode{} is na\"{i}ve type-directed 
    synthesis on the bijective lens combinators.  Our synthesis algorithm performs
    better than the na\"{i}ve approach and other string transformation systems,
    and our optimizations speed up the algorithm enough that all tasks become
    solvable.
  }
  \label{fig:synthesis-times}
\end{figure}

To generate specifications for Flash Fill, we generated input/output
specifications by generating random elements of the source language, and
running the lens on those elements to generate elements of the target language.
These were then fed to Flash Fill.

To generate specifications for FlashExtract, we extracted portions of strings
mapped in the generated lens either through an identity
transformation or through a previously synthesized lens, whereas strings that were
mapped through use of \ConstLens{} were considered boilerplate and so not
extracted.

As these tools were designed for a broader audience, they put less of a burden
on the user.  These tools only use input/output examples (for Flash
Fill), or marked text regions (for FlashExtract), as opposed
to \Optician{}'s use of regular expressions to constrain the format of
the input and output.  By using regular expressions,
\Optician{} is able to synthesize significantly more programs
than either existing tool.

Flash Fill and FlashExtract have two tasks: to determine how
the data is transformed, they must also infer the structure of the data, a
difficult job for complex formats.
In particular, neither Flash Fill nor FlashExtract was able to synthesize
transformations or extractions present under two iterations, a type of format
that is notoriously hard to infer.
These types of dual iterations are pervasive in Linux configuration
files, making Flash Fill and FlashExtract ill suited for many of the synthesis
tasks present in our test suite.

Furthermore, as unidirectional transformations, Flash Fill and FlashExtract have
a more expressive calculus.  To guarantee bidirectionality, our syntax must be
highly restrictive, providing a smaller search space to traverse.
